---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
params:
  agg.level: 'state'
  n.days.filter: 20
  web.version: FALSE
  extrap.date: '2020-01-26'
  count.start.date: '2020-03-01'
  end.data.date: '2020-09-05'
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

extrap.date <-  as.Date(params$extrap.date)
count.start.date <- as.Date(params$count.start.date)
end.data.date <- as.Date(params$end.data.date)

state.name2 <- c(state.name, 'District of Columbia','Puerto Rico', 'United States', 'New York City')
state.abb2 <- c(state.abb, 'DC','PR','US','NYC')


last.date.format <- 
  format(end.data.date, '%b %d, %Y')
```

```{r setup}
library(ExcessILI)
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(abind)
library(pbapply)
library(gsubfn)
library(dplyr)
library(RCurl)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(rjags)
library(INLA)
library(lme4)
#library(jsonlite)
set.seed(123)
source('./R/gen_pred_interval_inla.R')

```

## Population size data
https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic


 01 ... Under 1 year (includes not stated infant ages)
 02 ... 1 - 4 years
# 03 ... 5 - 14 years
# 04 ... 15 - 24 years
# 05 ... 25 - 34 years
# 06 ... 35 - 44 years
# 07 ... 45 - 54 years
# 08 ... 55 - 64 years
# 09 ... 65 - 74 years
# 10 ... 75 - 84 years
# 11 ... 85 years and over
# 12 ... Age not stated
```{r}
pop1 <- read.csv('./Data/SC-EST2020-alldata6.csv')

pop1$agec <- NA
pop1$agec[pop1$AGE>=0 & pop1$AGE<1] <- '01'
pop1$agec[pop1$AGE>=1 & pop1$AGE<5] <- '02'
pop1$agec[pop1$AGE>=5 & pop1$AGE<15] <- '03'
pop1$agec[pop1$AGE>=15 & pop1$AGE<25] <- '04'
pop1$agec[pop1$AGE>=25 & pop1$AGE<35] <- '05'
pop1$agec[pop1$AGE>=35 & pop1$AGE<45] <- '06' #category 85=85+; cat 999 is all combined
pop1$agec[pop1$AGE>=45 & pop1$AGE<55] <- '07' #category 85=85+; cat 999 is all combined
pop1$agec[pop1$AGE>=55 & pop1$AGE<65] <- '08' #category 85=85+; cat 999 is all combined
pop1$agec[pop1$AGE>=65 & pop1$AGE<75] <- '09' #category 85=85+; cat 999 is all combined
pop1$agec[pop1$AGE>=75 & pop1$AGE<85] <- '10' #category 85=85+; cat 999 is all combined
pop1$agec[pop1$AGE>=85 & pop1$AGE<110] <- '11' #category 85=85+; cat 999 is all combined

pop1$sex2 <- NA
pop1$sex2[pop1$SEX==1] <- 'M'
pop1$sex2[pop1$SEX==2]  <- 'F'
pop1$sex2[pop1$SEX==0]  <- 'ALL'

pop1$state.abb1 <- state.abb[match(pop1$NAME,state.name)]
pop1$state.abb1[pop1$NAME=='District of Columbia'] <- 'DC'

pop1$region <- NA
  pop1$region[pop1$state.abb1 %in% c('ME','VT','NH','MA','CT','RI','NY','NJ','PA')] <- 'Northeast'
  pop1$region[pop1$state.abb1 %in% c('DE','DC','FL','GA','MD','NC','SC','VA','WV','AL','MS','KY','TN','AK','LA','OK','TX')] <- 'South'
  pop1$region[pop1$state.abb1 %in% c('AZ','CA','ID','NM','MT','UT','WY','NV','AK','CA','HI','OR','WA')] <- 'West'
  pop1$region[pop1$state.abb1 %in% c('IN','IL','MI','OH','WI','IA','KS','MN','MO','NE','ND','SD')] <- 'Midwest'
  


pop1.m <- reshape2::melt(pop1[ pop1$ORIGIN==0 ,c('agec','region','RACE','sex2',"POPESTIMATE2010", "POPESTIMATE2011", "POPESTIMATE2012", "POPESTIMATE2013",  "POPESTIMATE2014", "POPESTIMATE2015", "POPESTIMATE2016", "POPESTIMATE2017", "POPESTIMATE2018", "POPESTIMATE2019" , "POPESTIMATE2020" )], id.vars=c('agec','sex2', 'RACE','region'))

pop2 <- reshape2::dcast(pop1.m, agec + sex2 +RACE +region+ variable ~., fun.aggregate = sum)

pop2$year <- as.numeric(gsub( 'POPESTIMATE','',as.character(pop2$variable)))

names(pop2) <- c('agec','SEX','RACE','region','variable','popsize','year')

pop2 <- pop2[,c('agec','SEX','RACE','region','popsize','year')]

pop2$qtr <- 1 #July 1

#pop2 <- pop2[pop2$state !='US',]

#This is a hacky fix because we only have pop estimates up until July 2020; assume it is constant until july 2021 for now
# fill_2021 <- pop2[pop2$year==2020,]
# fill_2021$year <- 2021
# 
# pop2 <- bind_rows(pop2, fill_2021)

empty_ts <- readRDS('./Data/Confidential/compiled_sex_age_race_qtr_region.rds')
empty_ts <- empty_ts[,c('agec','sex','region','race_recode','year','qtr')]
names(empty_ts) <- c('agec','sex','region','race','year','qtr')
  
names(pop2) <- tolower(names(pop2))

pop2 <- pop2[!is.na(pop2$agec),]
pop2 <- pop2[pop2$sex !='ALL',]


empty_ts <- empty_ts[empty_ts$agec !='12',]


empty_ts$qtr <- as.numeric(empty_ts$qtr)

empty_ts2 <- merge(empty_ts, pop2, by=c('agec','sex','race','region','year','qtr'), all=T)

empty_ts2$date[empty_ts2$qtr==3] <- (paste(empty_ts2$year[empty_ts2$qtr==3] ,'07','01', sep='-'))

empty_ts2$date[empty_ts2$qtr==4] <- (paste(empty_ts2$year[empty_ts2$qtr==4] ,'10','01', sep='-'))

empty_ts2$date[empty_ts2$qtr==1] <- (paste(empty_ts2$year[empty_ts2$qtr==1] ,'01','01', sep='-'))

empty_ts2$date[empty_ts2$qtr==2] <- (paste(empty_ts2$year[empty_ts2$qtr==2] ,'04','01', sep='-'))

empty_ts2$date <- as.Date(empty_ts2$date)
empty_ts2 <- empty_ts2[empty_ts2$date>='2014-07-01' ,] #UPDATE THIS!

empty_ts2 <- empty_ts2[empty_ts2$race !=6,]



filled_pop2 <- empty_ts2 %>%
      group_by(agec , sex,race, region) %>%
      arrange(agec , sex,race,region,year, qtr) %>%
          mutate(time=seq(1,n())) %>%
           mutate(pop.interpol=approx(time,popsize,time)$y)


filled_pop2 <- filled_pop2[,c('agec','sex','race','region','date','pop.interpol')]

# test1 <- filled_pop2[filled_pop2$state=='NY' & filled_pop2$age_group=="75-84 years",]
# plot(test1$week_end, test1$pop.interpol)
```

## Set up data for multilevel model

```{r}

ag3 <- readRDS('./Data/Confidential/compiled_sex_age_race_qtr_region.rds')
ag3$month <- NA
ag3$month[ag3$qtr==1] <- 1
ag3$month[ag3$qtr==2] <- 4
ag3$month[ag3$qtr==3] <- 7
ag3$month[ag3$qtr==4] <- 10
ag3$date <- as.Date(paste(ag3$year, ag3$month, '01', sep='-'))

ag3$qtr2 <- 0
ag3$qtr2[ag3$qtr==2] <- 1

ag3$qtr3 <- 0
ag3$qtr3[ag3$qtr==3] <- 1

ag3$qtr4 <- 0
ag3$qtr4[ag3$qtr==4] <- 1


#ag3 <- ag3[ag3$state!='US',]

ag3$time <- round(as.numeric(ag3$date - min(ag3$date))/30.3)

ag3$age_group <- as.factor(ag3$agec)
ag3$race_recode <- as.factor(ag3$race_recode)
ag3$sex <- as.factor(ag3$sex)

#ag3$state <- as.factor(ag3$state)

#mod.inla <- inla(all_cause ~  , family='poisson')

ag3$all_cause_pre <- ag3$N_deaths

ag3$all_cause_pre[ag3$date >= '2020-03-01'] <- NA
ag3$time_scale <- as.vector(scale(ag3$time))

# ag3.pre <- ag3[ag3$week_end < '2020-03-01',] %>%
#   group_by(state,age_group) %>%
#   summarize( log.ave.case = log(mean(all_cause)  ))
  
#ag3 <- merge( ag3, ag3.pre, by=c('state','age_group'))



ag3 <- merge(ag3, filled_pop2, by.x=c('sex','race_recode','agec','region', 'date'),by.y=c('sex','race','agec','region' ,'date'), all.x=T)


ag3$log_pop <- log(ag3$pop.interpol/100000)

ag3$subgroup_combo <- as.factor(paste(ag3$race_recode, ag3$agec, ag3$sex,ag3$region))
ag3$subgroup_combo2 <- ag3$subgroup_combo
ag3$subgroup_combo3 <- ag3$subgroup_combo
ag3$subgroup_combo4 <- ag3$subgroup_combo
ag3$subgroup_combo5 <- ag3$subgroup_combo

ag3$agec2 <- ag3$agec
ag3$agec3 <- ag3$agec
ag3$agec4 <- ag3$agec
ag3$agec5 <- ag3$agec

ag3$race_recode2 <- ag3$race_recode
ag3$race_recode3 <- ag3$race_recode
ag3$race_recode4 <- ag3$race_recode
ag3$race_recode5 <- ag3$race_recode

ag3$sex2 <- ag3$sex
ag3$sex3 <- ag3$sex

#ag3$state2 <- ag3$state
#ag3$state2 <- ag3$state

ag3 <- ag3[!is.na(ag3$pop.interpol),]

ag3$obs <- as.factor(1:nrow(ag3))

ag3$region <- as.factor(ag3$region)

#ag3 <- ag3[ag3$year<=2020,]
```

## Multilevel model

Simple GLM
```{r}
mod.glm1 <- glm(all_cause_pre ~ 
                1 + time_scale + qtr2 + qtr3 +qtr4 +
                  agec + race_recode + sex + region + subgroup_combo +
                  agec*(time_scale + qtr2 + qtr3 +qtr4) +
                  race_recode*(time_scale + qtr2 + qtr3 +qtr4) +
                  region*(time_scale + qtr2 + qtr3 +qtr4)  #+
                  #subgroup_combo*(time_scale + qtr2 + qtr3 +qtr4)
                  #sex*(time_scale + sin1 + sin2 + sin3 + cos1 +cos2+ cos3) + 
                     ,
                offset = log_pop,
                family='poisson', data=ag3)
summary(mod.glm1)
```


```{r, eval=F}
library(rstanarm)

mod.lme42 <- stan_glmer(all_cause_pre ~ 
                1 + time_scale +qtr2 + qtr3 +qtr4 +
                   sex  +(1+time_scale +qtr2 + qtr3 +qtr4|agec) + (1+time_scale +qtr2 + qtr3 +qtr4|race_recode), 
                offset = log_pop,
                data=ag3,
                family='poisson')

```


Model fit takes about 1 min
```{r}
# #RIDGE REGRESSION
#     #https://bitbucket.org/hrue/r-inla/src/0d9389b414979bcf3507c60f69a50488d18e4c6c/r-inla.org/examples/stacks/stacks.R?at=default
# 
# ag3.nomiss <- ag3[!is.na(ag3$region),]
# X <- model.matrix(~
#               -1 + time_scale + qtr2 + qtr3 +qtr4 +
#                   agec + race_recode + sex + subgroup_combo +
#                   agec*(time_scale + qtr2 + qtr3 +qtr4) +
#                   race_recode*(time_scale )  +
#                 #  sex*(time_scale + qtr2 + qtr3 +qtr4) +
#                   region*(time_scale + qtr2 + qtr3 +qtr4)  + 
#               subgroup_combo*(time_scale + qtr2 + qtr3 +qtr4)     
# ,
#                                  data=ag3.nomiss)
# 
#  first.beta<-paste0("f(beta1,", names(X)[1], "model='iid',hyper = param.beta,", "values=c(", paste(1:ncol(X), collapse=',') ,"))" )
#  
#    param.beta = list(prec = list(param = c(1.0e-3, 1.0e-3))) 
# 
#     next.betas <-  paste0("f(beta",2:ncol(X), names(X[-1]), ",copy='beta1', fixed=T)")
# 
#     next.betas<- paste(next.betas, collapse='+')  
#     all.betas<-paste(first.beta,next.betas , sep='+')   
# 
#           form1<- as.formula(paste0("all_cause_pre ~ 1+",  all.betas    ) )
#           
#           mod.df <- cbind.data.frame('all_cause_pre'=ag3.nomiss$all_cause_pre, X, 'log_pop'=ag3.nomiss$log_pop)
#           
#     mod1.full = inla(form1, data = mod.df, 
#                      control.predictor = list(compute=TRUE, link = 1), offset = log_pop,
#                      family='nbinomial')
# 
# summary(mod.inla2)
# 
# 
# preds.inla2 <- gen_pred_interval_inla(mod.inla2, ag3)
```

```{r}

mod.inla2 <- inla(all_cause_pre ~
              1 + time_scale + qtr2 + qtr3 +qtr4 +
                  agec + race_recode + sex +  subgroup_combo +
                  agec*(time_scale + qtr2 + qtr3 +qtr4) +
                  race_recode*(time_scale + qtr2 + qtr3 +qtr4)+
                  region*(time_scale + qtr2 + qtr3 +qtr4)

                  #sex*(time_scale + qtr2 + qtr3 +qtr4)
                               , offset = log_pop,
                family='poisson', data=ag3,
                control.compute=list(config = TRUE),
              control.fixed=list(mean=0, prec=1)
)
summary(mod.inla2)


preds.inla2 <- gen_pred_interval_inla(mod.inla2, ag3)
```



We can then take the results from the INLA and aggregate over different groups. For example, by age:
```{r}
preds.age <- preds.inla2$preds_covars.m %>%
  group_by(agec, date, variable) %>%
  summarise( 'N_deaths'=sum(N_deaths), 'pred'=sum(value) ) %>%
  ungroup() %>%
  group_by(agec, date) %>%
  summarise( 'N_deaths'=mean(N_deaths), 'pred_median'=median(pred), pred_lcl=quantile(pred,probs=0.025), pred_ucl=quantile(pred,probs=0.975) )


```

```{r, fig.width=8, fig.height=12}


 p1 <- ggplot(preds.age, aes(x=date, y=N_deaths , group=agec, col=agec)) +
    geom_line() +
    geom_line(aes(x=date, y=pred_median, group=agec), lty=2, col='black') +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
    geom_vline(xintercept=as.Date('2020-03-01'), col='gray', lty=2) +

        facet_wrap(~agec , scales='free') +
    xlim(as.Date('2014-01-01'), as.Date('2020-12-31')) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 

p1
```

or by age and race
```{r}
preds.age.race <- preds.inla2$preds_covars.m %>%
  group_by(agec,race_recode, date, variable) %>%
  summarise( 'N_deaths'=sum(N_deaths), 'pred'=sum(value) ) %>%
  ungroup() %>%
  group_by(agec, race_recode,date) %>%
  summarise( 'N_deaths'=mean(N_deaths), 'pred_median'=median(pred), pred_lcl=quantile(pred,probs=0.025), pred_ucl=quantile(pred,probs=0.975) )


```

```{r}
 p1 <- ggplot(preds.age.race[preds.age.race$agec=='08',], aes(x=date, y=N_deaths )) +
    geom_line() +
    geom_line(aes(x=date, y=pred_median, group=agec), lty=2, col='black') +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
    geom_vline(xintercept=as.Date('2020-03-01'), col='gray', lty=2) +

        facet_wrap(~race_recode , scales='free') +
    xlim(as.Date('2014-01-01'), as.Date('2020-12-31')) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 

p1
```