---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
params:
  agg.level: 'state'
  n.days.filter: 20
  web.version: FALSE
  extrap.date: '2020-01-26'
  count.start.date: '2020-03-01'
  end.data.date: '2020-09-05'
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

```

```{r setup}
library(ExcessILI)
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(patchwork)
library(abind)
library(gsubfn)
library(dplyr)
library(RCurl)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(rjags)
library(INLA)
library(lme4)
library(pbapply)
#library(jsonlite)
set.seed(123)

source('./R/process_va_region.R')
source('./R/process_nchs_region.R')
source('./R/inla_mod_grp.R')
source('./R/plot_ts_region.R')
source('./R/run_if_expired.R')
```



```{r, eval=F}
if (!require("devtools")) {
  install.packages("devtools")
}
devtools::install_github("weinbergerlab/ExcessILI")
#Test12343
```

Pulls in latest version of data from NCHS. can set update=F if you just want to use most local version available locally
```{r}
nchs.age.state <- runIfExpired('nchs_age_state', update=F,
 ~read.csv("https://data.cdc.gov/api/views/y5bj-9g5w/rows.csv?accessType=DOWNLOAD") 
 )

age1.region <- process_nchs_region(nchs.age.state)
```

VA data
```{r}
v1 <- read.csv('./Data/Confidential/aggdata_withregion.csv')
v2 <- process_va_region(v1)
```

Combine NCHS and VA data
```{r}

c1 <- bind_rows( v2 ,age1.region)

c1$one <- 1

max.start <- max(min(v2$week_end) , min(age1.region$week_end))
min.end <- min(max(v2$week_end) , max(age1.region$week_end))

c1 <- c1[c1$week_end>=max.start & c1$week_end<=min.end,]

c1 <- c1[order(c1$source, c1$age_group, c1$week_end),]

c1 <- c1[c1$week_end <= '2021-01-01',]

c1 <- c1[c1$region %in% c('Midwest','Northeast','South', 'West'), ]

```



```{r, fig.width=12, fig.height=5}

p1 <- plot_ts_region(source_data='NCHS')
p2 <- plot_ts_region(source_data='VA')
p2/p1


```

Try to overlay on same plot

```{r, fig.width=10, fig.height=10}

#Scale the time series
c1 <- c1 %>%
  group_by(age_group, region, source) %>%
  mutate(deaths_scale = (all_cause-mean(all_cause))/sd(all_cause) ) %>%
  ungroup()

p3 <- ggplot(c1 ,aes(x=week_end, y=deaths_scale, group=source, col=source) )  +
      geom_line() +
      ylab("Number of all-cause deaths") +
      xlab("Date") +
      theme_classic() +
      theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
      geom_hline(yintercept=0, col='gray', lty=2) +
        facet_wrap(region~age_group , scales='free') +
        theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 
p3

```

## run the model
```{r}
 statevar = "source"
 agevar='age_region'
 datevar = "week_end"
 outcome='all_cause'
 
c1$age_region <- paste(c1$age_group, c1$region, sep='_')
mod1 <- excessCases(c1, 
             statevar = statevar,
             agevar=agevar,
             datevar = datevar, 
             use.syndromes = c("all_cause"),
            extrapolation.date = as.Date("2020-03-01"), 
            sum.dates = as.Date("2020-03-01"),
            denom.var = "one",
            extend.epiyear =TRUE,
            time.res='week',
            model.type='poisson') 

```

## Interactive dashboard plot

```{r}
app1 <-dashboardPlot(mod1)
print(app1)
```


Extract output as arrays
 Can use this to calculate all sorts of custom things, like excess aggregated over different time periods. Some of the outputs contain draws from a Monte Carlo simulation, which can be used for propagating errors. For example, below we have sum.obs.ac gives the total deaths during the evaluation periods, and sum.pred.iter gives 10,000 values of the sum predicted value. These could be combined across strata
```{r}
ds <- mod1
dates1 <-
  ds[[1]][[1]][[1]]$date
  
sum.obs.ac <-
    excessExtract(ds = ds,
                syndrome = 'all_cause',
                extract.quantity = "sum.obs")

sum.pred.iter <-
    excessExtract(ds = ds,
                syndrome = 'all_cause',
                extract.quantity = "sum.pred.iter")

pred.iter <-
    excessExtract(ds = ds,
                syndrome = 'all_cause',
                extract.quantity = "pred.iter")

```

excess by age group and databse

```{r}
sum.obs.ac.tot <- apply(sum.obs.ac,c(2,3), sum)

rr.iter <- apply(sum.pred.iter, c(1), function(x) sum.obs.ac.tot/x)

rr.q <- t(apply(rr.iter,1, quantile, probs=c(0.025, 0.5, 0.975)))

```

 Sum across age groups, by database
Print the 95% CIs for the RR (overall observed/expected deaths by state)
```{r, fig.width=7, fig.height=7}

#pred.state.iter <- apply(sum.pred.iter,c(1,2), sum)

sum.obs.ac.state <- apply(sum.obs.ac, c(2,3), sum)

RR.state <- sapply(1:10000,function(x) sum.obs.ac.state/sum.pred.iter[x,,] , simplify='array')

RR.state.q <- round(apply(RR.state,c(1,2),quantile, probs=c(0.025,0.5,0.975)),2)

dimnames(RR.state.q)[[2]] <- dimnames(sum.obs.ac.state)[[1]]
dimnames(RR.state.q)[[3]] <- dimnames(sum.obs.ac.state)[[2]]


RR.state.q.m <- melt(RR.state.q)
RR.state.q.c <- dcast(RR.state.q.m, Var2+Var3 ~ Var1)

names(RR.state.q.c)[1:2] <- c('source', 'age_region') 

lab1 <- as.data.frame(matrix(unlist(str_split(RR.state.q.c$age_region, fixed("_"))), ncol=2, byrow=T))
names(lab1) <- c('age_group', 'region')
  
RR.state.q.c <- cbind.data.frame(RR.state.q.c, lab1)

p1 <-ggplot(RR.state.q.c, aes( y=`50%`, x=age_group, group=source, col=source)) + 
  geom_point(position = position_dodge(0.3))+
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0, position = position_dodge(0.3)) +
  theme_classic() +
      theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
      geom_hline(yintercept=1, col='gray', lty=2) +
      facet_wrap(~region ) 
p1
```



```{r}
excess_output = mod1
            statevar = "source"
             agevar='age_region'
             datevar = "date" 
             outcome='all_cause' 
             yaxis=statevar
             facet=agevar

  df_y <- 
    excess_output  %>% 
    pluck(outcome) %>% 
    map(~(
      transpose(.x) %>% 
        map(~(.x %>% bind_rows())) %>% .[c(datevar, "y")] %>% 
        map(~(.x %>% gather(statevar, y))) %>% 
        bind_cols()  %>% 
        set_names(c(statevar, datevar, "rm", "y")) %>% 
        dplyr::select(-rm)
    )) %>% 
    bind_rows(.id = agevar) 
  
  
```

rr_time_series
```{r}

rr.ls <- lapply(names(mod1[[1]]) , function(x){
  ds <- mod1[[1]][[x]]
  rr.nchs <- ds$NCHS$y/ds$NCHS$pred 
  rr.va <- ds$VA$y/ds$VA$pred 
  date <- ds$NCHS$date
  return(cbind.data.frame(date, rr.nchs, rr.va, 'age_group'=x))
})

rr <- bind_rows(rr.ls)
rr <- rr[rr$date>='2019-01-01',]

rr.m <- melt(rr, id.vars = c('age_group','date'))

p4 <- ggplot(rr.m ,aes(x=date, y=value, group=variable , col=variable ) )  +
      geom_line() +
      ylab("Observed.expected") +
      xlab("Date") +
      theme_classic() +
       ggtitle ('Observed/expected in va vs NCHS')+ 
      theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
      geom_hline(yintercept=1, col='gray', lty=2) +
        facet_wrap(~age_group ) +
        theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 
p4
```


## Population size data
https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic
```{r}
pop1 <- read.csv('./Data/SC-EST2020-alldata6.csv')

pop1$age_group <- NA
pop1$age_group[pop1$AGE>=0 & pop1$AGE<25] <- '0-25 years'
pop1$age_group[pop1$AGE>=25 & pop1$AGE<45] <- '25-44 years'
pop1$age_group[pop1$AGE>=45 & pop1$AGE<65] <- '45-64 years'
pop1$age_group[pop1$AGE>=65 & pop1$AGE<75] <- '65-74 years'
pop1$age_group[pop1$AGE>=75 & pop1$AGE<85] <- '75-84 years'
pop1$age_group[pop1$AGE>=85 & pop1$AGE<110] <- '85 years and older' #category 85=85+; cat 999 is all combined

pop1$state2 <- state.abb[match(pop1$NAME,state.name)]
pop1$state2[pop1$NAME=='District of Columbia'] <- 'DC'

pop1.m <- reshape2::melt(pop1[pop1$SEX==0 & pop1$ORIGIN==0 ,c('age_group','state2',"POPESTIMATE2010", "POPESTIMATE2011", "POPESTIMATE2012", "POPESTIMATE2013",  "POPESTIMATE2014", "POPESTIMATE2015", "POPESTIMATE2016", "POPESTIMATE2017", "POPESTIMATE2018", "POPESTIMATE2019" , "POPESTIMATE2020" )], id.vars=c('age_group','state2'))

pop2 <- reshape2::dcast(pop1.m, age_group + state2 + variable ~., fun.aggregate = sum)
pop2$year <- as.numeric(gsub( 'POPESTIMATE','',as.character(pop2$variable)))

names(pop2) <- c('age_group','state','variable','popsize','year')

pop2 <- pop2[,c('age_group','state','popsize','year')]

pop2$weekN <- 27 #July 1

pop2 <- pop2[pop2$state !='US',]

#This is a hacky fix because we only have pop estimates up until July 2020; assume it is constant until july 2021 for now
fill_2021 <- pop2[pop2$year==2020,]
fill_2021$year <- 2021

pop2 <- bind_rows(pop2, fill_2021)

empty_ts <- readRDS('./Data/cleaned_file.rds')
empty_ts <- empty_ts[,c('age_group','state','week_end')]
empty_ts <- empty_ts[empty_ts$state!='US',]

empty_ts$year <-year(empty_ts$week_end)
empty_ts$weekN <-week(empty_ts$week_end)

empty_ts2 <- merge(empty_ts, pop2, by=c('state','age_group','year','weekN'), all=T)

filled_pop2 <- empty_ts2 %>%
      group_by(age_group , state) %>%
      arrange(age_group, state, year, weekN) %>%
          mutate(time=seq(1,n())) %>%
           mutate(pop.interpol=approx(time,popsize,time)$y)

filled_pop2 <- filled_pop2[!is.na(filled_pop2$week_end),]

filled_pop2 <- filled_pop2[,c('age_group','state','week_end','pop.interpol')]

# test1 <- filled_pop2[filled_pop2$state=='NY' & filled_pop2$age_group=="75-84 years",]
# plot(test1$week_end, test1$pop.interpol)
```

## Set up data for INLA model

```{r}

ag3 <- c1

ag3 <- ag3 %>%
  group_by(region, age_group , source) %>%
  arrange(week_end) %>%
  mutate( time= row_number()) %>%
  ungroup()

ag3$sin1 <- sin(2*pi*ag3$time/52.1775)
ag3$sin2 <- sin(2*pi*ag3$time*2/52.1775)
ag3$sin3 <- sin(2*pi*ag3$time*3/52.1775)

ag3$cos1 <- cos(2*pi*ag3$time/52.1775)
ag3$cos2 <- cos(2*pi*ag3$time*2/52.1775)
ag3$cos3 <- cos(2*pi*ag3$time*3/52.1775)
 
ag3$age_group <- as.factor(ag3$age_group)

ag3$region <- as.factor(ag3$region)

#mod.inla <- inla(all_cause ~  , family='poisson')

ag3$all_cause_pre <- ag3$all_cause

ag3$all_cause_pre[ag3$week_end >= '2020-03-01'] <- NA

ag3$time_scale <- as.vector(scale(ag3$time))

ag3$year <- year(ag3$week_end)
ag3$weekN <- week(ag3$week_end)

#ag3 <- merge(ag3, filled_pop2, by=c('state','age_group', 'week_end'), all.x=T)


#ag3$log_pop <- log(ag3$pop.interpol)

# ag3$state_ag_combo <- as.factor(paste(ag3$state, ag3$age_group))
# ag3$state_ag_combo2 <- ag3$state_ag_combo
# ag3$state_ag_combo3 <- ag3$state_ag_combo

ag3$age_group2 <- ag3$age_group
ag3$age_group3 <- ag3$age_group

# ag3$state2 <- ag3$state
# ag3$state3 <- ag3$state

ag3$quarter <- quarter(ag3$week_end)
ag3$age_region_source <- paste(ag3$region, ag3$age_group, ag3$source, sep='_')

#ag3 <- ag3[!is.na(ag3$pop.interpol),]

#ag3 <- ag3[ag3$year<=2020,]
```


## Simple GLM, in INLA. Fit separately by group

```{r}

mod1 <- pblapply(unique(ag3$age_region_source), inla_mod_group , ds.in=ag3)

preds_total <- bind_rows(lapply(mod1,'[[','preds_total' ))
preds_total_qtr <- bind_rows(lapply(mod1,'[[','preds_qtr' ))
preds_total_wk <- bind_rows(lapply(mod1,'[[','preds_week' ))

#Splits out the labels we need and add as column
lab_split <- function(ds, var_to_split){
  ds <- ds%>%
    mutate(region = sapply(strsplit(as.character(ds[,var_to_split]), "_"),"[", 1),
           age_group= sapply(strsplit(as.character(ds[,var_to_split]), "_"),"[", 2), 
           source =sapply(strsplit(as.character(ds[,var_to_split]), "_"),"[", 3))
  return(ds)
}

preds_total_wk <- lab_split(preds_total_wk, 'age_region_source')
preds_total_qtr <- lab_split(preds_total_qtr, 'age_region_source')
preds_total <- lab_split(preds_total, 'age_region_source')


```


```{r}

plot_ts_region <- function( source_data='NCHS'){
  
  p1 <- ggplot(preds_total_wk[preds_total_wk$source==source_data,], aes(x=week_end, y=all_cause, group=region, col=region)) +
    geom_line() +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
    facet_wrap(region~age_group , scales='free', nrow=1) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 
  return(p1)
}
```

RRs
```{r}
preds_total$RR_mean <- preds_total$all_cause/preds_total$pred_mean
preds_total$RR_lcl <- preds_total$all_cause/preds_total$pred_lcl
preds_total$RR_ucl <- preds_total$all_cause/preds_total$pred_ucl


p1 <-ggplot(preds_total, aes( y=RR_mean, x=age_group, group=source, col=source)) + 
  geom_point(position = position_dodge(0.3))+
  geom_errorbar(aes(ymin=RR_lcl, ymax=RR_ucl), width=0, position = position_dodge(0.3)) +
  theme_classic() +
      theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
      geom_hline(yintercept=1, col='gray', lty=2) +
      facet_wrap(~region ) 
p1
```






## Multilevel model




Model fit takes about 1 min
```{r}

mod.inla2 <- inla(all_cause_pre ~ 
                1 + time_scale + sin1 + sin2 + sin3 + cos1 +cos2+ cos3 +
                  age_group + state + state_ag_combo +
                  age_group*(time_scale + sin1 + sin2 + sin3 + cos1 +cos2+ cos3) +
                       state*(time_scale + sin1 + sin2 + sin3 + cos1 +cos2+ cos3) +
                        
                #Random slopes for the time trends and sin1, cos1 effects
                f(state_ag_combo, time_scale,model='iid') +
                  
                f(state_ag_combo2, sin1,model='iid') +
                f(state_ag_combo3, cos1,model='iid') 
                    ,
                                offset = log_pop,
                family='poisson', data=ag3,
                control.compute=list(config = TRUE))
summary(mod.inla2)

pred.sample.list <- inla.posterior.sample(n=500, mod.inla2, seed=123)

post.labels<-dimnames(pred.sample.list[[1]]$latent)[[1]]

posterior.samples<- sapply(pred.sample.list, '[[', 'latent') #should be 'joint' instead of 'latent' https://www.ucl.ac.uk/population-health-sciences/sites/population-health-sciences/files/inla_baio.pdf

preds.select<-grep('Predictor',post.labels )

posterior.preds<-exp(posterior.samples[preds.select,]) #lambda

preds.summary <- t(apply(posterior.preds,1, quantile, probs=c(0.025,0.5,0.975)))

ag3.preds <- cbind.data.frame(preds.summary,ag3)

summary.list <- list( 'ag3.preds'=ag3.preds)

saveRDS(summary.list,'./Data/INLA_results.rds')



```

```{r, fig.width=8, fig.height=12}
summary.list <- readRDS('./Data/INLA_results.rds')

summary.spl <- split(summary.list[[1]], paste(summary.list[[1]]$state, summary.list[[1]]$age_group ))

states <- sapply(summary.spl, function(x) unique(x$state))
ages <- sapply(summary.spl, function(x) unique(x$age_group))


lapply(unique(ages), function(y){
  summary.spl.age <- summary.spl[ages==y]
  par(mfrow=c(6,3), mar=c(2,2,1,1))
  lapply(summary.spl.age, function(x){
    age_state_lab <- paste(unique(x$age_group),unique(x$state), sep=' ')
    
    plot(x$week_end, x$all_cause, type='l', bty='l', main=age_state_lab)
    points(x$week_end, x$'50%', type='l', col='red', lty=2)
    points(x$week_end, x$'50%', type='l', col='red', lty=2)
    abline(v=as.Date('2020-03-01'), lty=2, col='gray')
    
    })

})

```


